Viscous Erosion in Porous Media

### 2018-03-20 docs/methods/erosion_methods.tex
NM

To compare with this analytical prediction, we aim to measure the opening angles formed in our numerical simulations. However, our simulations lose accuracy just in the final moments as the body is vanishing due to the shear stress blowing up (in accordance with the scaling laws discussed in Section \ref{sec:scaling}). To circumvent this problem, we could restart our simulation with a scaled-up version of the the near-final  body as the initial condition and iterate the process until convergence. A simpler solution, though, is to remove the effect of shrinking by modifying the normal velocity $\Vn$ to have zero mean,
\begin{align*}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align*}
This modification allows erosion to sculpt the body as it normally would but artificially keeps the area fixed. We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process at equally-spaced times. Notice the body obtains a similar profile as observed previously, but with fixed area. Importantly, more numerical accuracy is now available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) shows a zoom of the body's rear stagnation point, demonstrating that the near-corner is actually smooth on a fine scale, in accordance with our numerical smoothing. 
Likewise, Fig.~\ref{fixed_intface}(b) shows the stress distribution on the evolving body, revealing that stress becomes more uniform as time proceeds. Interestingly, the slight rise in shear near the front/rear corners that was seen in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that the effect was due to loss of accuracy.
Here, time is normalized by the same vanishing time $t_f = 1.79E-2$ as in Fig.~\ref{shrink_intface}. Curiously, about 3 multiples of $t_f$ are required to reach an approximate steady-state, which contrasts with previous studies of high-Reynolds-number erosion, in which a satisfactory steady-state was usually reached before the body vanished~\cite{moo-ris-chi-zha-she2013}.

@@

The vorticity field not only allows us to visualize regions of high shear, but also allows us to identify preferred channels with most intense flow. Figure~\ref{fig:06bodies} shows that the flow is concentrated primarily in a single channel that runs horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels follow (e.g.~between bodies 1 \&2 or 3 \& 5 or 5 \& 6). Since the pressure drop across the cell is fixed, the overall flow rate increases as the bodies erode. This can be seen by the overall vorticity magnitude growing in time. By the final frame shown, body 4 has vanished completely and our method handles this event seamlessly.

>>>>

To compare with this analytical prediction, we aim to measure the opening angles formed in our numerical simulations. However, our simulations lose accuracy just in the final moments as the body is vanishing due to the shear stress blowing up (in accordance with the scaling laws discussed in Section \ref{sec:scaling}). To circumvent this problem, we could restart our simulation with a scaled-up version of the the near-final  body as the initial condition and iterate the process until convergence. A simpler solution, though, is to remove the effect of shrinking by modifying the normal velocity $\Vn$ to have zero mean,
\begin{align*}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align*}
This modification allows erosion to sculpt the body as it normally would but artificially keeps the area fixed. We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process at equally-spaced times. Notice the body obtains a similar profile as observed previously, but with fixed area. Importantly, more numerical accuracy is now available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) shows a zoom of the body's rear stagnation point, demonstrating that the near-corner is actually smooth on a fine scale, in accordance with our numerical smoothing. 
Likewise, Fig.~\ref{fixed_intface}(b) shows the stress distribution on the evolving body, revealing that stress becomes more uniform as time proceeds. Interestingly, the slight rise in shear near the front/rear corners that was seen in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that the effect was due to loss of accuracy.
Here, time is normalized by the same vanishing time $t_f = 1.79 \times 10^{-2}$ as in Fig.~\ref{shrink_intface}. Curiously, about 3 multiples of $t_f$ are required to reach an approximate steady-state, which contrasts with previous studies of high-Reynolds-number erosion, in which a satisfactory steady-state was usually reached before the body vanished~\cite{moo-ris-chi-zha-she2013}.

@@

The vorticity field not only allows us to visualize regions of high shear, but also allows us to identify preferred channels with most intense flow. Figure~\ref{fig:06bodies} shows that the flow is concentrated primarily in a single channel that runs horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels follow (e.g.~between bodies 1 \& 2 or 3 \& 5 or 5 \& 6). Since the pressure drop across the cell is fixed, the overall flow rate increases as the bodies erode. This can be seen by the overall vorticity magnitude growing in time. By the final frame shown, body 4 has vanished completely and our method handles this event seamlessly.
